Verification of Stencil Computations through Spatial Specifications

### 2016-07-03 paper.tex
Sexier name
Separate coalesced region to make understanding easier
Fix math typos in inference & better partial order on matrices
added some words about the corpus and scripts to pull in the results and stats
remove infer-output-july3.txt and replace it with the next results file...

\title{Abstract Spatial Specification of Stencil Computations}
\authorinfo{Dominic Orchard \and Mistral Contrastin
\and Matthew Danish \and Andrew Rice}{University of Cambridge}{}

@@

Informally, the procedure proceeds as follows. First, relative
indices are turned into unit regions in $n$-dimensions. For
example, for the five point stencil there are five $1 \times 1$ regions,
called \emph{spans}, which are essentially $0$-dimensional:
%
\begin{center}
\begin{tikzpicture}
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (1,1) rectangle (1.5,1.5);
\draw[thick] (1.5,1) rectangle (2,1.5);
\draw[thick] (1,0.5) rectangle (1.5,1);
\draw[thick] (0.5,1) rectangle (1,1.5);
\draw[thick] (1,1.5) rectangle (1.5,2);
\end{tikzpicture}
\end{center}
%
Each dimension is traversed, coalescing consecutive spans
which vary in $m$-dimensions into coalesced spans which vary in
$m+1$-dimensions. In this example, the $0$-dimensional spans
 become $1$-dimensional spans (rows/columns):
%
\begin{align*}
\begin{array}{cc}
\begin{tikzpicture}
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (0.5,1) rectangle (2,1.5);
\draw[thick] (1,0.5) rectangle (1.5,1);
\draw[thick] (1,1.5) rectangle (1.5,2);
\end{tikzpicture}
\qquad
&
\qquad
\begin{tikzpicture}
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (1.5,1) rectangle (2,1.5);
\draw[thick] (0.5,1) rectangle (1,1.5);
\draw[thick] (1,0.5) rectangle (1.5,2);
\end{tikzpicture}
\\
\textit{dimension 1} \qquad & \qquad \textit{dimension 2}
\end{array}
\end{align*}
%
Any spans that are contained with any other span are deleted,
leaving a minimal set of spans (which may overlap, but none of which
fully contains another).
\begin{center}
\begin{tikzpicture}
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (1,0.5) rectangle (1.5,2);
\node at (2,1.22) {+};
\draw[thick] (2.5,1) rectangle (4,1.5);
%\node at (2.75,1.22) {\textnormal{\small{0,0}}};
\end{tikzpicture}
\end{center}
%
The procedure is then iterated till a fixed point is reached. In this
example this is reached in the first step. These final spans
go on to form the basis of the spatial stencil specification, where
each span is a region product \texttt{*} and multiple spans are
combined with the region sum \texttt{+}.

@@

\section{Evaluation}
\label{sec:evaluation}

>>>>

\include{results}
\include{corpussize}

\title{Verification of Stencil Computations through Spatial Specifications}
\authorinfo{Dominic Orchard \and Mistral Contrastin
\and Matthew Danish \and Andrew Rice}{University of Cambridge}{}

@@

Informally, the procedure proceeds as follows. First, relative
indices are turned into unit regions in $n$-dimensions. For
example, for the five point stencil there are five $1 \times 1$ regions,
called \emph{spans}, which are essentially $0$-dimensional:
%
\begin{center}
\begin{tikzpicture}
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (1,1) rectangle (1.5,1.5);
\draw[thick] (1.5,1) rectangle (2,1.5);
\draw[thick] (1,0.5) rectangle (1.5,1);
\draw[thick] (0.5,1) rectangle (1,1.5);
\draw[thick] (1,1.5) rectangle (1.5,2);
\end{tikzpicture}
\end{center}
%
Each dimension is traversed, coalescing consecutive spans
which vary in $m$-dimensions into coalesced spans which vary in
$m+1$-dimensions. In this example, the $0$-dimensional spans
 become $1$-dimensional spans (rows/columns):
%
\begin{align*}
\begin{array}{cc}
\begin{tikzpicture}
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (0.5,1) rectangle (2,1.5);
\draw[thick] (1,0.40) rectangle (1.5,0.90);
\draw[thick] (1,1.60) rectangle (1.5,2.10);
\end{tikzpicture}
\qquad
&
\qquad
\begin{tikzpicture}
%\draw[step=0.5cm,lightgray,very thin] (0.5,0.5) grid (2,2);
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (1.6,1) rectangle (2.1,1.5);
\draw[thick] (0.4,1) rectangle (0.9,1.5);
\draw[thick] (1,0.5) rectangle (1.5,2);
\end{tikzpicture}
\\
\textit{dimension 1} \qquad & \qquad \textit{dimension 2}
\end{array}
\end{align*}
%
Any spans that are contained with any other span are deleted,
leaving a minimal set of spans (which may overlap, but none of which
fully contains another).
\begin{center}
\begin{tikzpicture}
%\node at (1.25,1.22) {\textnormal{\small{0,0}}};
\draw[thick] (1,0.5) rectangle (1.5,2);
\node at (2,1.22) {+};
\draw[thick] (2.5,1) rectangle (4,1.5);
%\node at (2.75,1.22) {\textnormal{\small{0,0}}};
\end{tikzpicture}
\end{center}
%
The procedure is then iterated till a fixed point is reached. In this
example this is reached in the first step. These final spans
go on to form the basis of the spatial stencil specification, where
each span is a region product \texttt{*} and multiple spans are
combined with the region sum \texttt{+}.

@@

\section{Evaluation}
\label{sec:evaluation}
We now present an evaluation of the stencil language. We first confirm
that stencils are common in scientific code and further that our
stencil language can express a great many of these.

\subsection{Software Corpus}
\begin{figure}
\begin{tabular}{|l|l|l|}
\hline
Package       & Files & Lines of code \\
\hline
UM            & \umFiles  & \umLoC       \\
E3MG          & \ethreemgeaFiles   & \ethreemgeaLoC        \\
BLAS          & \blasFiles   & \blasLoC        \\
Hybrid4       & \hybridfourFiles    & \hybridfourLoC          \\
GEOS-Chem     & \geoschemFiles   & \geoschemLoC       \\
Navier        & \navierFiles     & \navierLoC           \\
CP            & \computationalphysicstwoFiles    & \computationalphysicstwoLoC           \\
\hline
\textbf{Total}& \overallFiles  & \overallLoC     \\
\hline 
\end{tabular}
\caption{Summary of software packages used for evaluation\label{fig:corpus}}
\end{figure}

Figure \ref{fig:corpus} shows summary statistics of the software packages used in our evaluation. In total we analysed 1.1 million lines of code from 7 packages.

\textbf{The Unified Model}\footnote{\url{http://www.metoffice.gov.uk/research/modelling-systems/unified-model}} is a weather forecasting and climate modelling tool developed by the Met Office in the United Kingdom. It is used by research organisations and meterological services around the world. We use the development branch (trunk) of the model. The Met Office runs a comprehensive code quality system incorporating dedicated commiters (we counted 11) for particular parts of the model. We counted 120 additional contributors whose submissions are reviewed and tested before being accepted into the code base.

% e3mg-ea
\textbf{E3MG} (An Energy-Environment-Economy (E3) Model at the Global Level) is a macroeconomic model used for asssessment of environmental policy~\cite{RePEc:aen:journl:2006se-a12}. This model was developed by Cambridge Econometrics, an independent consultancy company.

\textbf{BLAS} (Basic Linear Algebra Subprograms)\footnote{http://www.netlib.org/blas/} provides efficient and portable routines for vector and matrix operations. These routines feature in many other libraries (including LAPACK). We used version 3.6.0.

\textbf{Hybrid4} is a vegetation and biomass model for simulating carbon, water and nitrogen flows~\cite{GBC:GBC635}.

\textbf{GEOS-Chem}\footnote{\url{http://acmg.seas.harvard.edu/geos/}} is a three-dimensional model of tropospheric chemistry developed at Harvard and used by approximately 70 universities and research institutions world-wide. We examined version 10-01.

\textbf{Navier} is an implementation of a Navier-Stokes fluid dynamics computation~\cite{griebel1997numerical}.

\textbf{CP} consists of the example code from the book ``Computational Physics''~\cite{nicholas2006computational} introducing numerical techniques and their application to modern physics problems such as fields, waves, statistical mechanics and quantum mechanics.

% atLeast: number of stencils with an atLeast modifier
% atMost: number of stencils with an atMost modifier
% boundedBoth: number of stencils with both atLeast and atMost modifier
%% atLeast == atMost == boundedBoth - whenever we infer one we infer the other but users might write just one

% backward: number of backward stencils
% forward: number of forward stencils
% centereed: number of centered stencils

% justReflexive: number of only reflexive stencils
% irreflexive: number of irreflexive stencils
% readOnce: number of readOnce stencils

% depthX: number of stencils with depth X
% dimsX: number of stencils with dimensions=X (check this!)

% lexFailed: number of files we failed to lex
% parseFailed: number of files we failed to parse
% lexOrParseFailed: number files we failed to lex and parse
% parseOk: number of files parsed successfully
%% lexFailed + parseFailed = lexOrParseFailed
%% lexOrParseFailed + parseOk = number of files in source repo

% tickAssign: number of array subscripts with a LHS which are all neighbour indices or constants
% tickAssignSuccess: number of tickAssign's we generated some specification for
%% tickAssign < tickAssignSuccess
